---
title: "Peaky Blinders Network Analysis"
author: "Matteo Larrode"
date: "2024-12-05"
output:
  pdf_document: default
  html_document: default
---

# Peaky Blinders Network Analysis (Part 2)

## Peaky Blinders Network

```{r message=FALSE, warning=FALSE}
library(igraph)
library(tidyverse)
library(gridExtra)
library(ergm)
library(intergraph)
library(network)
library(knitr)
```

First, let's load the adjacency matrix of character interactions created previously. 
```{r}
peaky_df <- readRDS("data/cooccurrence_df.rds")

peaky_adj_mat <- as.matrix(peaky_df)
# no self-links
diag(peaky_adj_mat) <- 0
```

Now, we can create the network from the adjacency matrix:
```{r}
peaky_network <- graph_from_adjacency_matrix(peaky_adj_mat,
  mode = "undirected",
  weighted = TRUE
)

peaky_network <- set_vertex_attr(
  peaky_network,
  "name_edited",
  value = str_to_title(str_replace_all(V(peaky_network)$name, "_", " "))
)

peaky_network
```

And we can produce an initial plot!

```{r}
set.seed(10)
layout2 <- layout_with_kk(peaky_network)
layout4 <- layout_with_lgl(peaky_network)

plot.igraph(peaky_network,
  edge.color = "gray",
  edge.curved = .1,
  edge.width = 1 + E(peaky_network)$weight / 15,
  vertex.size = 3 + degree(peaky_network) / 3,
  vertex.frame.color = "#555555",
  vertex.label = V(peaky_network)$name_edited,
  vertex.color = "#FBD87F",
  vertex.label.color = "black",
  vertex.label.cex = 1 + betweenness(peaky_network, weights = NA) / 1000,
  margin = c(0, 0, 0, 0),
  asp = 0,
  layout = layout4,
  main = "Network of Peaky Blinders (Seasons 1-6)"
)
```

## Assortativity and communities

### Distinct subgroups, modularity and assortativity

Let's divide the network into three mutually exclusive subgroups depending of their city/country of origin, and their sex.

```{r}
# Orign
birmingham <- c(
  "Finn", "Arthur", "Thomas", "Freddie", "Danny", "Polly", "Ada",
  "Charlie", "John", "Curly", "Jeremiah", "Lizzie", "Michael", "Esme",
  "Karl", "Linda", "Aberama", "Bonnie", "Isiah", "Younger", "Erasmus",
  "Anna", "Moss", "Ruben", "Frances", "Barney", "Billy Kimber",
  "Johnny Dogs", "Jessie Eden", "Mrs Connors"
)

london <- c(
  "Churchill", "May", "Alfie", "Sabini", "Mitford", "Oswald Mosley", "Sabini",
  "Goliath", "Cyril", "Mr Levitt"
)

new_york <- c(
  "Luca Changretta", "Jack Nelson", "Vicente", "Gina", "Mrs Changretta"
)

ireland <- c(
  "Grace", "Campbell", "Hughes", "Laura Mckee", "Billy Grade", "Hayden Stagg"
)

russia <- c(
  "Tatiana", "Romanov", "Izabella"
)

# Gender
male <- c(
  "Finn", "Arthur", "Thomas", "Churchill", "Freddie", "Danny",
  "Charlie", "John", "Campbell", "Curly", "Younger", "Jeremiah",
  "Erasmus", "Michael", "Moss", "Karl", "Alfie", "Sabini",
  "Ruben", "Vicente", "Romanov", "Hughes", "Darby", "Aberama",
  "Bonnie", "Isiah", "Goliath", "Cyril", "Chang", "Barney",
  "Billy Kimber", "Johnny Dogs", "Luca Changretta", "Jimmy Mccavern",
  "Oswald Mosley", "Jack Nelson", "Mr Levitt", "Billy Grade",
  "Hayden Stagg"
)

female <- c(
  "Polly", "Ada", "Grace", "May", "Lizzie", "Esme", "Anna",
  "Izabella", "Frances", "Linda", "Tatiana", "Gina", "Mitford",
  "Mrs Connors", "Laura Mckee", "Evadne Barwell", "Mrs Changretta",
  "Jessie Eden"
)

# Create dataset
chr_variables <- data.frame(
  name_edited = V(peaky_network)$name_edited
) |>
  mutate(
    origin = case_when(
      name_edited %in% birmingham ~ "Birmingham",
      name_edited %in% london ~ "London",
      name_edited %in% new_york ~ "New York",
      name_edited %in% ireland ~ "Ireland",
      name_edited %in% russia ~ "Russia",
      TRUE ~ "Other"
    ),
    origin_colour = case_when(
      origin == "Birmingham" ~ "#AAF7FF",
      origin == "London" ~ "#FFE099",
      origin == "New York" ~ "#CCBFFF",
      origin == "Ireland" ~ "#B2FF8C",
      origin == "Russia" ~ "#F76D5E",
      origin == "Other" ~ "#F2DACD"
    ),
    gender = case_when(
      name_edited %in% male ~ "Male",
      name_edited %in% female ~ "Female",
      TRUE ~ "Unknown"
    ),
    gender_colour = case_when(
      gender == "Male" ~ "#FFFBA2",
      gender == "Female" ~ "#40B0A6",
      TRUE ~ "grey"
    )
  )
```

These character attributes are added to the network.

```{r}
peaky_network <- peaky_network |>
  set_vertex_attr("origin", value = chr_variables$origin) |>
  set_vertex_attr("origin_colour", value = chr_variables$origin_colour) |>
  set_vertex_attr("gender", value = chr_variables$gender) |>
  set_vertex_attr("gender_colour", value = chr_variables$gender_colour)


gender_membership <- as.integer(as.factor(V(peaky_network)$gender))
origin_membership <- as.integer(as.factor(V(peaky_network)$origin))
```

We can now visualise the different subgroups in the network by colouring its vertices.

```{r}
plot.igraph(peaky_network,
  edge.color = "gray",
  edge.curved = .1,
  edge.width = 1 + E(peaky_network)$weight / 15,
  vertex.size = 3 + degree(peaky_network) / 3,
  vertex.frame.color = "#555555",
  vertex.label = V(peaky_network)$name_edited,
  vertex.color = V(peaky_network)$gender_colour,
  vertex.label.color = "black",
  vertex.label.cex = 1 + betweenness(peaky_network, weights = NA) / 1000,
  margin = c(0, 0, 0, 0),
  asp = 0,
  layout = layout4,
  main = "Network of Peaky Blinders (Seasons 1-6)"
)

legend(
  x = -1.15, y = 0.9,
  legend = c("Male", "Female"),
  col = c("black", "black"),
  pt.bg = c("#FFFBA2", "#40B0A6"), # Background color of the points
  pch = 21, # Type of points
  pt.cex = 3, # Size of the points
  bty = "n", # No box
  text.col = "black"
)

text(0.7, -1,
  paste0("Modularity: ", round(modularity(peaky_network, membership = gender_membership), 3)),
  cex = 1.2,
  col = "black",
  font = 2
)
```

```{r}
plot.igraph(peaky_network,
  edge.color = "gray",
  edge.curved = .1,
  edge.width = 1 + E(peaky_network)$weight / 15,
  vertex.size = 3 + degree(peaky_network) / 3,
  vertex.frame.color = "#555555",
  vertex.label = V(peaky_network)$name_edited,
  vertex.color = V(peaky_network)$origin_colour,
  vertex.label.color = "black",
  vertex.label.cex = 1 + betweenness(peaky_network, weights = NA) / 1000,
  margin = c(0, 0, 0, 0),
  asp = 0,
  layout = layout4,
  main = "Network of Peaky Blinders (Seasons 1-6)"
)

legend(
  x = -1.2, y = 1.3,
  legend = levels(as.factor(V(peaky_network)$origin)),
  col = c("black", "black"),
  pt.bg = levels(as.factor(V(peaky_network)$origin_colour)), # Background color of the points
  pch = 21, # Type of points
  pt.cex = 3, # Size of the points
  bty = "n", # No box
  text.col = "black"
)

text(0.7, -1,
  paste0("Modularity: ", round(modularity(peaky_network, membership = origin_membership), 3)),
  cex = 1.2,
  col = "black",
  font = 2
)
```

Now, we can calculate a modularity statistic using these subgroups, and interpret what the results imply for assortativity. 

```{r}
paste0("Modularity with respect to gender: ", modularity(peaky_network, membership = gender_membership))
paste0("Modularity with respect to origin: ", modularity(peaky_network, membership = origin_membership))
```

The modularity statistics for gender (0.0087) and origin (0.0139) are very close to zero. For instance, the proportion of character interactions within the same origin is a mere 0.0139 percentage points higher than what would be expected if all ties were formed at random.

In the context of Peaky Blinders, these modularity statistics indicate a very low level of assortativity based on gender and origin, i.e., interactions between characters are not markedly influenced by these attributes. Characters in the series therefore tend to interact across gender and origin lines, supporting a narrative where character interactions are primarily driven by plot needs and personal relationships rather than by their demographic backgrounds. 

### Continuous variable and assortativity 

Let's now calculate normalised closeness for all the nodes, and study assortativity with respect to this continuous variable. 

```{r}
paste0(assortativity(peaky_network, closeness(peaky_network, normalized = TRUE)))
```

There is a slight positive assortativity by closeness: the correlation between the closeness of linked vertices is 0.12 which is positive, albeit fairly weak. Characters which are closer on average to others in the network tend to interact more often amongst each other than with more peripheral characters, suggesting a subtle trend towards interaction among similarly influential characters within the narrative structure.

### Community detection algorithm

Having failed to detect a notable community, either by gender or origin earlier does not signify that there are no underlying subgroups influencing interactions between the characters of Peaky Blinders. Community detection algorithms can help detect these hidden communities.

To reduce subjectivity, multiple algorithms are compared: two node-moving algorithms based on modularity optimisation (Fast-greedy and Louvain) and one that minimises the expected description length of a random walker trajectory (InfoMAP).

#### Fast-greedy algorithm

```{r}
set.seed(123)

communities_fg <- cluster_fast_greedy(peaky_network)
table(membership(communities_fg))
```

The algorithm found four communities, let's visualise them.

```{r}
plot.igraph(peaky_network,
  edge.color = "gray",
  edge.curved = .1,
  edge.width = 1 + E(peaky_network)$weight / 15,
  vertex.size = 3 + degree(peaky_network) / 3,
  vertex.frame.color = "#555555",
  vertex.label = V(peaky_network)$name_edited,
  vertex.color = membership(communities_fg),
  vertex.label.color = "black",
  vertex.label.cex = 1 + betweenness(peaky_network, weights = NA) / 1000,
  margin = c(0, 0, 0, 0),
  asp = 0,
  layout = layout4,
)

text(0.7, -1,
  paste0("Modularity: ", round(modularity(peaky_network, membership = membership(communities_fg)), 3)),
  cex = 1.2,
  col = "black",
  font = 2
)
```
#### Louvain algorithm

```{r}
set.seed(10)

communities_lou <- cluster_louvain(peaky_network)
table(membership(communities_lou))
```

The algorithm found agin four communities, but with different group sizes. Let's visualise these new groups.

```{r}
plot.igraph(peaky_network,
  edge.color = "gray",
  edge.curved = .1,
  edge.width = 1 + E(peaky_network)$weight / 15,
  vertex.size = 3 + degree(peaky_network) / 3,
  vertex.frame.color = "#555555",
  vertex.label = V(peaky_network)$name_edited,
  vertex.color = membership(communities_lou),
  vertex.label.color = "black",
  vertex.label.cex = 1 + betweenness(peaky_network, weights = NA) / 1000,
  margin = c(0, 0, 0, 0),
  asp = 0,
  layout = layout4,
)

text(0.7, -1,
  paste0("Modularity: ", round(modularity(peaky_network, membership = membership(communities_lou)), 3)),
  cex = 1.2,
  col = "black",
  font = 2
)
```

A confusion matrix can help understand the differences between the two methods:

```{r}
table(membership(communities_fg), membership(communities_lou))
```
Both node moving algorithms (Fast-Greedy and Louvain) find 4 communities. 27 of the 28 members of Community 3 detected by the Fast Greedy method are consistently placed in the same community by the Louvain method, suggesting that these nodes likely form a tightly-knit or clearly defined cluster within the network, which both algorithms recognize and consistently categorize. Similarly, 10 nodes out of eleven from Fast Greedy's Community 2 were placed together by the Louvain algorithm. 

Nevertheless, the confusion matrix shows some differences between the two algorithms’ assignments. For instance, for Community 1 of Fast Greedy, the nodes are spread across four different communities in the Louvain method. This indicates that the cohesion within this group as perceived by the Fast Greedy method is not as strong under the criteria used by the Louvain algorithm. 

#### InfoMAP community detection algorithm

Rather than optimising modularity, InfoMAP minimises the expected description length of a random walker trajectory. In other words, the partition into communities is done so that walks should spend longer time within communities than between.

```{r}
communities_infomap_unweighted <- cluster_infomap(peaky_network, e.weights = NULL)
communities_infomap <- cluster_infomap(peaky_network)
table(membership(communities_infomap_unweighted))
table(membership(communities_infomap))
```

Interestingly, community detection with InfoMAP (weighted and unweighted) results in a single community. This is an indication that the flow of interactions between characters does not significantly bottleneck at any point that would otherwise suggest the presence of distinct groups.

When contrasted with the outputs from the previous algorithms, which identified multiple communities, the result from InfoMAP might indicate that the network is overall tightly-knit but contains subtler structures that are more sensitive to the parameters and methods of node-moving algorithms.

#### Compare Louvain algorithm and grouping by origin

```{r}
sort(membership(communities_lou))

table(V(peaky_network)$origin, membership(communities_lou))
```

Cross-tabulating the communities detected by the algorithm and subgroups yields interesting results. Although modularity alone had hinted that origin might not be a decisive factor for tie formation, there are some overlaps with the community attribution made by the Louvain algorithm. 

- Community 1 predominantly consists of characters from Birmingham (19 out of 31), suggesting that this community could represent the central group of characters based in Birmingham, likely reflecting the primary setting and central figures Peaky Blinders. Unsurprinsingly, all men of the Shelby family are in this community (Thomas, Arthur, Finn, John), as well as their trusted officers (Danny, Curly, Johnny Dogs)

- The other community mostly constituted of characters originating from Birmingham is Community 2 (9 out of 11). Interestingly, this second Birmingham-related community includes the most influential women in the Shelby family, such as Polly, the matriarch (along with her lover Aberama). In fact, from the Shelby siblings, all men are in Community 1, and the only woman, Ada, is isolated in Community 2 (along with her husband, Freddie, and son, Karl). These findings suggests that gender may indeed have a role in the creation of ties, along with family ties. 

- 93% of the characters from Birmingham are in either Community 1 or 2. Community 3 and Community 4 appear to be smaller and more diverse in terms of character origin. Community 4 has a slight concentration of characters from London (3 out of 6). However, this community seems to be constituted along narrative lines. Indeed, five of its members (Alfie and his dog Cyril, Sabini, Darby and Luca Changretta) are major antagonists of Peaky Blinders. 

- Finally, Community 3 is the most diverse, including characters from all origins, and possibly hinting at more internationally connected narratives or character interactions. For example, although Michael originates from Birmingham, he marries Gina and moves to New York, where she is from, to conducts business with her uncle Jack Nelson, therefore connecting two continents through his interactions. 

Overall, the Louvain community detection algorithm has uncovered communities in the narrative of Peaky Blinders, hidden when only considering single attributes at a time, but seemingly driven by interactions based on a set of factors including place of origin, gender, family or business. 

## Small-world and scale-free networks

### Scale-free networks

- **Scale-free network**: In a scale-free network, most nodes have only a few connections, while a small number of nodes - often referred to as "hubs" - have a much large number of connections. A scale-free network is characterised by its highly uneven distribution of connections among the nodes. Specifically, this distribution follows a follows a power law, meaning that the probability of observing an node with 'x' connections decays as x increases: typically in the form P(x) = Cx^-α, with normalisation constant C and power law exponent α.

The power law is ubiquitous in nature and society, let's observe whether it characterises the interactions of Peaky Blinders characters.

First, degree distribution:
```{r}
degree_data <- degree_distribution(peaky_network, cumulative = FALSE)
cumulative_degree_data <- degree_distribution(peaky_network, cumulative = TRUE)

degree_df <- data.frame(
  degree = seq_along(degree_data) - 1,
  probability = degree_data
)

cumulative_degree_df <- data.frame(
  degree = seq_along(cumulative_degree_data) - 1,
  probability = cumulative_degree_data
)

p1 <- ggplot(degree_df, aes(x = degree, y = probability)) +
  geom_line() +
  geom_point() +
  labs(title = "Degree Distribution", x = "Degree", y = "p(degree)") +
  theme_minimal()

p2 <- ggplot(cumulative_degree_df, aes(x = degree, y = probability)) +
  geom_line() +
  geom_point() +
  labs(title = "Cumulative Degree Distribution", x = "Degree", y = "p(degree)>x") +
  theme_minimal()
```

```{r}
# Remove zero probability & degree for log scale
cumulative_degree_log_df <- cumulative_degree_df[cumulative_degree_df$probability > 0 & cumulative_degree_df$degree > 0, ]

log_log_cumulative_plot <- ggplot(cumulative_degree_log_df, aes(x = degree, y = probability)) +
  geom_point() +
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  labs(title = "Cumulative Degree Distribution \n(Log-Log)", x = "Degree (log scale)", y = "p(degree)>x (log scale)") +
  theme_minimal()

# Combining the plots side by side
grid.arrange(p1, p2, log_log_cumulative_plot, ncol = 3)
```

The first two plots hint at a potential power law distribution of degree. When picking a random node in the network, it has the highest probability of having a degree between 0 and 10, with a maximum at 3. Conversely, only a few nodes have especially high degrees (one at 54, and four others at 30 or above).

The log-log plot of the cumulative degree distribution is a key visualization for analyzing the scale-free characteristics of a network. Although, the beginning part of the cumulative degree distribution curve does seem to follow a rough linear pattern in the log-log scale, suggesting a power-law, there appears to be more variation and a potential drop-off towards the right side. This could indicate that while the network does exhibit some level of scale-free characteristics, it does not strictly follow a power-law distribution across all degrees, which could be due to the relatively small size of the network.

Using the degrees of the nodes, we can estimate and report the α exponent of the power law.

```{r}
degree_values <- degree(peaky_network, mode="in")

# Fit the power law
power_law_peaky <- fit_power_law(degree_values) 

power_law_peaky
```
- α Value: The α value for is estimated as 5.287987. This is quite high, as typical α values are usually observed between 2 and 3 for networks that follow a power-law distribution, suggesting that the network has a steeper drop-off at higher degrees.

- xmin: The xmin value is estimated at 28, indicating that the power-law behavior applies to degrees greater than or equal to 28. This means that the distribution of node degrees becomes scale-free starting at a relatively high threshold, suggesting that most nodes below this degree do not follow the power-law pattern.

- KS.p Value: The Kolmogorov-Smirnov p-value is very high (0.9999997), which means there is a strong likelihood that the observed degree distribution is consistent with a power-law distribution, at least for degrees higher than xmin. With such a high p-value, we cannot reject the hypothesis that the degree distribution follows a power-law.

Overall, the analysis suggests that the Peaky Blinders network is consistent with a power-law distribution for higher degree values (degrees >= 28), with a steep decline in node connectivity. This indicates a network where a few nodes are highly connected, which vastly outnumber the less connected nodes, a typical characteristic of scale-free networks. However, the unusually high alpha value points to an even more pronounced inequality in node connectivity than is typically seen in many real-world networks.

The scale-free characteristics of this network can also be assessed with the 20-80 rule. 

```{r}
# ---- UNWEIGHTED DEGREE ----
# 20% of the 57 nodes is 12 nodes
total_connections <- sum(degree_values)

# Sort nodes by degree in descending order and get the sum of degrees for the top 20%
top_degrees <- sum(sort(degree_values, decreasing = TRUE)[1:12])

# Calculate the percentage of connections these top nodes hold
percentage_connections_held_by_top <- (top_degrees / total_connections) * 100

cat("Percentage of connections held by the top 20%: ", round(percentage_connections_held_by_top, 2), "%\n")

# ---- WEIGHTED DEGREE ----
weighted_degree_values <- strength(peaky_network)
total_weight <- sum(weighted_degree_values)

# Sort nodes by weighted degree in descending order and get the sum of weighted degrees for the top 20%
top_weights <- sum(sort(weighted_degree_values, decreasing = TRUE)[1:12])

# Calculate the percentage of connections these top nodes hold
percentage_weights_held_by_top <- (top_weights / total_weight) * 100

cat("Percentage of weighted degrees held by the top 20%: ", round(percentage_weights_held_by_top, 2), "%\n")
```

A significant portion of the network, nearly half of the nodes (49.28%), is required to account for 80% of the connections. While this concentration of connections among fewer than half of the nodes indicates some level of centralization, it does not meet the typical characteristics of a scale-free network, where we would expect around 20% of the nodes to control 80% of the connections. This network's connectivity pattern is more centralized than in a fully random network, where connections would be more evenly distributed across all nodes, but it does not exhibit the extreme centralization often associated with scale-free networks, which confirms the interpretation of the figures above.

Interestingly, when considering the weighted degrees, the top 20% of nodes hold 75.95% of the total connection weights, suggesting a more significant concentration of influence within a small subset of the network.

Several mechanisms can lead to the formation of scale-free networks, most of them are usually driven by some form of feedback loop. For example, preferential attachment - when new nodes prefer to attach to nodes that are already well-connected - and duplication are two of such mechanisms. A network of TV series character interactions is a special case because its interactions are solely driven by choices of the screenwriter to advance the plot, not by decisions made by the nodes themselves. 

If scale-free characteristics were observed in this network, they would caused by the need to advance the plot around only a few characters (e.g. the Shelby family), and if side characters did interact amongst each other. Had the series had additional seasons, with its network growing further, it is expected that its scale-free proporty would have been enhanced, as side characters or those deceased, and antagonists from previous seasons would have stopped appearing in the show and therefore been unable to interact at all with new characters.

### Small-world pheomenon

- **Small-world phenomenon**: The small-world phenomenon refers to the property of a network wherein most nodes can be reached from every other node by a small number of steps, despite the network being very large. It is often quantified with a low average shortest path length, and is most likely to be exhibited by networks whose edges are mostly formed at random.

```{r}
# ---- WITH ALL NODES ----
# Calculate all shortest paths in the network
distances <- distances(peaky_network, weights = NA)
shortest_distances <- distances[upper.tri(distances)]

# Calculate the average shortest path length
average_path_length <- mean_distance(peaky_network, unconnected = FALSE, weights = NA)

# Create a data frame of the original shortest path distances
hist_data <- data.frame(
  PathLengths = shortest_distances,
  Network = "Original")

# ---- REMOVING THOMAS ----
peaky_network_modified <- delete_vertices(peaky_network, c("thomas"))
distances_modified <- distances(peaky_network_modified, weights = NA)
shortest_distances_modified <- distances_modified[upper.tri(distances_modified)]

# Recalculate the average shortest path length in the modified network
average_path_length_modified <- mean_distance(peaky_network_modified, unconnected = TRUE, weights = NA)

hist_data_modified <- data.frame(
  PathLengths = shortest_distances_modified,
  Network = "Without Thomas") |> 
  filter(is.finite(PathLengths))

# Combine the data frames for original and modified networks
combined_hist_data <- rbind(hist_data, hist_data_modified)

# Plotting histograms side by side
distances_hist <- ggplot(combined_hist_data, aes(x = PathLengths, fill = Network)) +
  geom_bar(color = "black", position = "dodge") +
  scale_x_continuous(breaks = seq(0, 4, by = 1)) +
  labs(x = "Shortest Path Length", y = "Frequency") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "None")+
  annotate("text",
           x = 3.6, y = 1073,
           label = paste0("Average Path Length: ", round(average_path_length, 2)),
           fontface = "bold",
           col = "#f7766d",
           size = 7) +
  annotate("text",
           x = 2.975, y = 883,
           label = paste0("Average Path Length: ", round(average_path_length_modified, 2),"\n(after removing Thomas)"),
           fontface = "bold",
           col = "#02bfc4",
           hjust = 0,
           size = 7)

distances_hist
```

The average path length of 1.83 and this figure suggest that the network, which includes 57 nodes, does exhibit small-world properties. Thomas Shelby, main character of Peaky Blinders (with a degree of 54), is the hub that connects most of the characters together. Indeed, he occupies most of the screen time of the show, and therefore interacts with almost all characters. Interestingly, the small-world properties of the network persist even after removing Thomas from the network, with the average path length only increasing to 2.03. This indicates that other characters, including Arthur and Polly, also act as hubs, allowing every character to reach all the others in four or less steps.

Had Peaky Blinders kept going after its 6 seasons, the only situation where its network would have stopped exhibiting these small-world characteristics is if the narrative structure of the series had completely changed, and stopped following a restricted set of characters. In this situation, the network could have lost its main hubs and become less closely connected.

## Exponential random graph modelling (ERGM)

Unlike many approaches, ERGMs adequately captures network dependencies,  whereby the existence of one link in the network can influence the existence of others. By including node- and edge-level characteristics, and structural features, ERGMs enable  adequate hypothesis testing and simulations.

### Hypotheses

I formulate three hypotheses regarding the Peaky Blinders network to be tested using an ERGM:

1. **H1 (Homophily): Characters from the same origin are more likely to interact with each other than with those from different origins**

2. **H2: Characters from Birmingham have more ties than those from other origins.**

Indeed, the Shelby family, focal set of characters in the series, all originate from Birmingham, and therefore concentrate a large number of ties.

3. **H3: (Clustering):  There is a tendency for triadic closure, i.e., characters are more likely to interact if they share mutual connections.**

Indeed, characters who appear together in shared scenes or storylines are likely to develop interconnected relationships, thus forming triangles within the network of character interactions.


### ERGM's

```{r echo=FALSE, message=FALSE}
# Converting peaky_network to network format
peaky_network2 <- asNetwork(peaky_network)

# Fit the ERGM models
# ---- ONLY EDGES ----
peaky_ERGM1 <- ergm(peaky_network2 ~ edges)

# ---- ADDING TERM FOR HOMOPHILY ----
peaky_ERGM2 <- ergm(peaky_network2 ~ edges + nodematch("origin"))

# ---- ADDING DUMMY VARIABLES FOR ORIGIN OF THE NODE ----
peaky_ERGM3 <- ergm(peaky_network2 ~ edges + nodematch("origin") + nodefactor("origin"))

# ---- ADDING TRIADIC CLOSURE ----
peaky_ERGM4 <-  ergm(peaky_network2 ~ edges + nodematch("origin") + nodefactor("origin") + gwesp(0.25,T) + degree(3))
```

Now let's intepret the results of these models, and test our different hypotheses. 

```{r}
summary(peaky_ERGM1)

paste0("Density of the Peaky Blinders network: ", round(edge_density(peaky_network), 3))
paste0("Probability of a tie: ", plogis(coef(peaky_ERGM1)[["edges"]]))
```

Converting the negative estimate (-1.29) for the 'edges' variable in the baseline model (model 1) to the probability that two nodes are connected with a tie leads to the density of the network, 21.6%. 

```{r}
# H1: Homophily
summary(peaky_ERGM2)

paste0("The odds of a within-origin tie are ", round(exp(coef(peaky_ERGM2)[["nodematch.origin"]]), 3), " greater than that of a between-origin tie.")
paste0("Probability of a tie between characters of different origins: :", plogis(coef(peaky_ERGM2)[["edges"]]))
paste0("Probability of a tie between characters of the same origins: :", plogis(coef(peaky_ERGM2)[["edges"]] + coef(peaky_ERGM2)[["nodematch.origin"]]))
```

Model 2 shows initial strong support for H1, with the odds of a within-origin tie 3.59 times (e^1.2767) higher than the odds of a between-origin tie. The 14.3% probability of a tie formed between characters of different origins increases to 37.4% when the characters share their origin. 

```{r}
# H2: Characters from Birmingham are more likely to share ties with other nodes
summary(peaky_ERGM3)

h2_df <- data.frame(
  log_odds = coef(peaky_ERGM3)[3:7]
  ) |> 
  mutate(
    odds = exp(log_odds)
  )

kable(h2_df)
```

Beyond providing further evidence for H1, with the homophily variable still displaying a positive and statistically significant estimate after controlling for the origin of the node, model 3 suppors H2. Compared to those from Birmingham, characters from all other locations show reduced odds of having ties, as displayed in the table above. 

```{r}
# H3: Transitivity
summary(peaky_ERGM4)
```

The global clustering coeﬃcient of the network is 0.46, meaning that there is 46% chance, which indicates a moderate tendency in this network to form closed triangles. This hypothesis is supported by results of model 4, with a positive and statistically significant estimate for the term for triadic closure.

Let us check differences in model fits between model 3 and model 4, which includes structural variables:

```{r}
anova(peaky_ERGM3, peaky_ERGM4)
```

### Goodness-of-fit analysis

To check the goodness of fit of the last, most elaborate model, a simulation will be carried out. After generating 100 networks from model 4, I will study how observed characteristics of the real network compare with the simulated ones. If the model is a good fit to the data, the simulated networks should look very much like the network we observed with respect to those network statistics. 

```{r}
set.seed(100)

peaky_ERGM4_GOF <- gof(peaky_ERGM4) 
```

```{r}
plot(peaky_ERGM4_GOF)
```

The goodness-of-fit diagnostics for Model 4 indicate a generally strong alignment between the model's simulations and the observed network data. The box plots in the top-left panel demonstrate that most observed network statistics, including edges and triadic closure, closely align with the median of the simulations, although there are slight deviations in node-specific characteristics like origin. The top-right panel shows an excellent match in the distribution of geodesic distances, suggesting the model accurately captures how connectivity is distributed across the network. In the bottom panel, the observed degree distribution matches well with the simulations for lower degrees, but shows some discrepancies at higher degrees. Overall, these diagnostics suggest that Model 4 effectively models the network structure and dynamics of the character interactions with strengths in depicting overall connectivity and triadic closure, though it could be further refined to better capture attributes associated with specific node origins and higher degrees.